---
title: "240131-cluster-x-betas"
format: html
---

## Here we go:

```{r echo = FALSE}
library(tidyverse)
library(ggplot2)
library(DBI)
library(RSQLite)
library(here)
library(glue)

# Loading files:
sim_name <- "240131-cluster-x"
results_folder <- here("Results", "Database-extracts", sim_name)

data_raw <- readRDS(file = here(results_folder, "betas.RDS")) 
data <- data_raw |> tibble()
cn <- colnames(data)
```

## Model selection: Hausman or Fixed
- Hausman test does very well in identifying false positives

```{r echo = FALSE}
data_select <- data |> 
  filter(model == "Hausman") |> 
  mutate(random = !is.na(est) |> as.integer())

for(myStudySize in c(50, 200, 1000)){
  for(myICC in c(0.05, 0.2)){
  plot <- data_select |> 
    filter(study_sample_size_train == myStudySize, ICC == myICC, random) |> 
    ggplot(mapping =aes(x = pred_icc)) +
    geom_bar() +
    facet_grid(rows = vars(n_studies), cols = vars(b_w_ratio)) +
    labs(title = "Number of times Hausman test select random intercept model", subtitle = glue("Study sample size: {myStudySize}, ICC: {myICC}"))
  
  print(plot)
  
  }
}



```

```{r echo = FALSE}
data_select <- data |> 
  filter(model == "Hausman") |> 
  mutate(random = !is.na(est) |> as.integer())

for(myStudySize in c(50, 200, 1000)){
  for(myICC in c(0.05, 0.2)){
  plot <- data_select |> 
    filter(study_sample_size_train == myStudySize, ICC == myICC, random) |> 
    ggplot(mapping =aes(x = pred_icc)) +
    geom_bar() +
    facet_grid(rows = vars(n_studies), cols = vars(b_w_ratio)) +
    labs(title = "Number of times Hausman test select random intercept model", subtitle = glue("Study sample size: {myStudySize}, ICC: {myICC}"))
  
  print(plot)
  
  }
}



```

## Error in Beta

```{r echo = FALSE}
for(myStudySize in c(50, 200, 1000)){
  for(myICC in c(0.05, 0.2)){
  plot <- data |> mutate(plot_value = beta_mean_error) |> 
    filter(study_sample_size_train == myStudySize, ICC == myICC) |> 
    ProfacSims:::box_plot_cluster_x() +
    ylab("Error in beta") +
    labs(title = "Error in beta estimate", subtitle = glue("Study sample size: {myStudySize}, ICC: {myICC}"))
  
  print(plot)
  
  }
}



```
## Root mean squared Error in Beta

```{r echo = FALSE}
for(myStudySize in c(50, 200, 1000)){
  for(myICC in c(0.05, 0.2)){
  plot <- data |> mutate(plot_value = beta_rmse) |> 
    filter(study_sample_size_train == myStudySize, ICC == myICC) |> 
    ProfacSims:::box_plot_cluster_x() +
    ylab("Root mean squared error for beta")
    labs(title = "Error in beta estimate", subtitle = glue("Study sample size: {myStudySize}, ICC: {myICC}"))
  
  print(plot)
  
  }
}
